Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-21 cav2017/cav2017.tex
some reductions

There are two sources of error due to numeric representation; the numeric error introduced by 
the fixed point numbers used to model our plant, i.e, the plant dynamics $A_d$, $B_d$ and the states; 
and the numeric error introduced by the digital controller, which uses an analogue-to-digital converter to read
the states as a fixed-point number. In this section we formally outline the errors introduced by
fixed point representation of numbers; in further sections we will differentiate between these
two errors by referring to them as the \emph{plant error} and the \emph{digital controller error}.

Let $\mathbb{R} \langle I,F \rangle$ be a fixed point domain with $I$ bits representing the integer part and $F$ bits
representing the decimal part. 
The following approximation errors will arise:
\begin{enumerate}
\item {\bf Truncation:}
Let $\mathcal{F}\langle I,F \rangle(\cdot) : \mathbb{R} \rightarrow \mathbb{R}\langle I,F \rangle, \tilde x=\mathcal{F}\langle I,F \rangle(x) = x-\delta : \delta=x\ \%_F\  c\tilde x$, where $\%_F$ is the modulus operation performed on the last bit of the representation ($2^{-F}$).
The number $\delta$ corresponds to the truncation error of the representation and it will propagate across operations.
\item {\bf Rounding:}
%% Every time an operation is performed on a number in the $\mathbb{R} \langle I,F,E \rangle$ domain, an error may be
%% introduced due to truncation of the result.
Let us define $c\langle i,f \rangle \in \mathbb{R}\langle I,F\rangle$, where $i \in [-2^{I-1}\ 2^{I-1}],f \in [0 1)$ are its assigned values. The smallest number that can be represented for any given exponent is $c_{min}=2^{-F}$.
The following errors appear in simple operations:
\begin{enumerate}
\item Addition/Substraction: in the case of fixed point arithmetic of numbers with the same domain, there are no rounding errors, otherwise the maximum error is given by the domain of the result $c\langle \cdot, F_1\rangle \pm c\langle \cdot , F_2\rangle= c\langle \cdot , F_3\rangle + \delta_3 :|\delta_3| \leq 2^{-F_3} \wedge F_3 \leq F_1,F_2$ 
\item Multiplication: $c\langle \cdot, F_1\rangle * c\langle \cdot,F_2\rangle=c\langle \cdot,F_3\rangle + \delta_3 : |\delta_3| \leq 2^{-F_3} \wedge F_3 \leq F_1,F_2$ 
\end{enumerate}
The bounds obtained for division are left to the reader as it will not be applied in this paper.

At this point it is worth mentioning that not all systems truncate equally. In the definition above $\delta$ may be positive
or negative, and even this decision can be made dependent on the sign of $x$. Common cases in computer systems are:
round downwards (\emph{ie} to $-\infty$), round upwards (\emph{ie} to $+\infty$), and round to zero.
Rounding errors are cumulative which means the overall error will statistically increase with every operation, thus algorithms
performing fewer operations can be more precise in this respect than iterative ones. 
\item {\bf Overflow:}
  %Another source of error introduced by Finite Word Length representations is that of overflow.
  The finite word length set a
maximum representable number in the domain ($\pm 2^{I-1}+1-2^{-F}$), which means that any number outside this range
cannot be accurately represented by the domain (\emph{eg} $\delta=x-2^{I-1}+1-2^{-F}$). In the case of floating point, a special
case for infinity has been introduced, which addresses some of this issues, but once the error has gone to infinity, the numbers
become meaningless. It is the responsibility of software designers and tools to account for overflows and either supply saturation
functions to prevent them or raise alarms in their presence. In the case of verification, the latter is seen as a safety problem.
\end{enumerate}

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Fixed word length effects on matrix multiplications}
\label{sec:cof_fwl}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Thus far we have presented a verification procedure for a digitalized plant in a continuous space (\emph{ie} the
range of the quatizers is still $\mathbb{R}$. The next step we must reaize is that the controller and the observer
exist in a digital program, hence their domain is $\mathbb{R}\langle I,F\rangle$. This has two separate effects
on the model. If we recall equation \eqref{eq:to_of}, the matrices of the observer and the controller  are not necessarily those of the plant.
In fact we have $\mat{A},\mat{B} \in \mathbb{R}^*$, $\mat{A}_o,\mat{B}_o,\mat{K}_d, \mat{L}_d \in \mathbb{R}\langle I,F\rangle^*$
Where $*$ represents the appropriate dimension for each matrix.
Unfortunately this means that although combined matrices such as $\mat{A}_t$ are in the reals, they suffer from
the fixed word restrictions in their composing matrices.
The second effect corresponds to the rounding errors in the calculation of $\hat{x}_{k+1}, u_k$. These will introduce a new source of quantiztion noise which we may denote $\vec{\nu}_3 \in [-q_3\ \ q_3]$.
The estimation of $\hat{\vec{x}}$ corresponds to the sum of three matrix-vector multiplications. Each of these has a coefficient-wise error bound $c_{min}$. Hence the overall $\overline{\delta}_{\hat{x}}=3nc_{min}$. The error in $\vec{u}$ corresponds of the multiplication of $\mat{K}_d\hat{\vec{x}}$, hence $q_3=\overline{\delta}_u=nc_{min}+\sum_{k=1}^nc_k\overline{\delta}_{\hat{x}}=(3|\mat{K}|_1+1)n2^{-F}$

>>>>

There are two sources of error due to numeric representation; the numeric error introduced by 
the fixed point numbers used to model our plant, i.e, the plant dynamics $A_d$, $B_d$ and the states; 
and the numeric error introduced by the digital controller, which uses an analogue-to-digital converter to read
the states as a fixed-point number and performs operations in this domain. In this section we formally outline the errors introduced by
fixed point representation of numbers; in further sections we will differentiate between these
two errors by referring to them as the \emph{plant error} and the \emph{digital controller error}.

Let $\mathbb{R} \langle I,F \rangle$ be a fixed point domain with $I$ bits representing the integer part and $F$ bits
representing the decimal part. 
The following approximation errors will arise:
\begin{enumerate}
\item {\bf Truncation:}
Let $\mathcal{F}\langle I,F \rangle(\cdot) : \mathbb{R} \rightarrow \mathbb{R}\langle I,F \rangle, \tilde x=\mathcal{F}\langle I,F \rangle(x) = x-\delta : \delta=x\ \%_F\  c\tilde x$, where $\%_F$ is the modulus operation performed on the last bit of the representation ($2^{-F}$).
The number $\delta$ corresponds to the truncation error of the representation and it will propagate across operations.
\item {\bf Rounding:}
%% Every time an operation is performed on a number in the $\mathbb{R} \langle I,F,E \rangle$ domain, an error may be
%% introduced due to truncation of the result.
Let us define $c\langle i,f \rangle \in \mathbb{R}\langle I,F\rangle$, where $i \in [-2^{I-1}\ 2^{I-1}],f \in [0 1)$ are its assigned values. The smallest number that can be represented for any given exponent is $c_{min}=2^{-F}$.
The following errors appear in simple operations:
\begin{enumerate}
\item Addition/Substraction: in the case of fixed point arithmetic of numbers with the same domain, there are no rounding errors, otherwise the maximum error is given by the domain of the result $c\langle \cdot, F_1\rangle \pm c\langle \cdot , F_2\rangle= c\langle \cdot , F_3\rangle + \delta_3 :|\delta_3| \leq 2^{-F_3} \wedge F_3 \leq F_1,F_2$ 
\item Multiplication: $c\langle \cdot, F_1\rangle * c\langle \cdot,F_2\rangle=c\langle \cdot,F_3\rangle + \delta_3 : |\delta_3| \leq 2^{-F_3} \wedge F_3 \leq F_1,F_2$ 
\end{enumerate}
The bounds obtained for division are left to the reader as they will not be applied in this paper.

\item {\bf Overflow:}
  %Another source of error introduced by Finite Word Length representations is that of overflow.
  The finite word length set a
maximum representable number in the domain ($\pm 2^{I-1}+1-2^{-F}$), which means that any number outside this range
cannot be represented by the domain (\emph{eg} $\delta=x-2^{I-1}+1-2^{-F}$). Our tool currently assumes no saturation
function and verifies that no state overflows.
\end{enumerate}

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Fixed word length effects on matrix multiplications}
\label{sec:cof_fwl}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Thus far we have presented a verification procedure for a digitalized plant in a continuous space (\emph{ie} the
range of the quatizers is still $\mathbb{R}$. The next step we must reaize is that the controller and the observer
exist in a digital program, hence their domain is $\mathbb{R}\langle I,F\rangle$. This has two separate effects
on the model. If we recall equation \eqref{eq:to_of}, the matrices of the observer and the controller  are not necessarily those of the plant.
In fact we have $\mat{A},\mat{B},\vec{x} \in \mathbb{R}^*$, $\mat{K}_d,\vec{u} \in \mathbb{R}\langle I,F\rangle^*$
Where $*$ represents the appropriate dimension for each matrix/vector.
Unfortunately this means that although combined matrices such as $\mat{A}_t$ are in the reals, they suffer from
the fixed word restrictions in their composing matrices.
The second effect corresponds to the rounding errors in the calculation of $\hat{x}_{k+1}, u_k$. These will introduce a new source of quantiztion noise which we may denote $\vec{\nu}_3 \in [-q_3\ \ q_3]$.
The estimation of $\hat{\vec{x}}$ corresponds to the sum of three matrix-vector multiplications. Each of these has a coefficient-wise error bound $c_{min}$. Hence the overall $\overline{\delta}_{\hat{x}}=3nc_{min}$. The error in $\vec{u}$ corresponds of the multiplication of $\mat{K}_d\hat{\vec{x}}$, hence $q_3=\overline{\delta}_u=nc_{min}+\sum_{k=1}^nc_k\overline{\delta}_{\hat{x}}=(3|\mat{K}|_1+1)n2^{-F}$
